% q0008.tex — No traces left behind (Basic Concepts / Mathematical Economics)
\question{
  id        = {0008},
  title     = {No Traces Left Behind},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Micro},
  subtopic  = {Mathematical Economics / Basic Concepts},
  type      = {objective},
  statement = {%
    In the nation of Fermatland, inflation follows an unconventional model.
    After the adoption of a new currency, a product originally priced at $p$
    in year~1 has value $p^t$ in year~$t$. There are 100 companies offering
    the same product, with company~$i$ having an initial price of~$i$
    (for $1 \leq i \leq 100$).

    Each year, the government provides a subsidy exactly equal to the price of
    one randomly selected company's product for that year. The subsidy must
    be spent entirely within the year and cannot be saved.

    You wonder whether it is ever possible to buy \emph{two} products in a
    single year using the subsidy exactly --- that is, with no money left over.

    How many years can this happen?

    \textit{Hint:} Fermat's Last Theorem states that there are no positive
    integer solutions to the equation $a^n + b^n = c^n$ for $n \geq 3$.
  },
  choices   = {%
    \item 0
    \item 1
    \item 2
    \item 100
    \item Forever
  },
  answer    = {C},
  solution  = {%
    In year $t$, company $i$ charges $i^t$. The subsidy equals some $k^t$
    (for a randomly chosen company $k$). We need $a^t + b^t = k^t$ for
    some $a, b, k \in \{1, \ldots, 100\}$.

    \begin{itemize}
      \item \textbf{Year $t=1$:} We need $a + b = k$. Many solutions exist,
            e.g.\ $3 + 5 = 8$. \textbf{Possible.}
      \item \textbf{Year $t=2$:} We need $a^2 + b^2 = k^2$ (Pythagorean
            triples). Solutions exist, e.g.\ $3^2 + 4^2 = 5^2 = 25$, and
            $3, 4, 5 \leq 100$. \textbf{Possible.}
      \item \textbf{Year $t \geq 3$:} By Fermat's Last Theorem, there are
            \emph{no} positive integer solutions to $a^t + b^t = c^t$.
            \textbf{Impossible.}
    \end{itemize}

    Therefore, it is possible in exactly \textbf{2 years} ($t=1$ and $t=2$).
    Answer: \textbf{(C)}.
  },
}
