% q0014.tex — Double it and give it to the next person
\question{
  id        = {0014},
  title     = {Double It and Give It to the Next Person},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Micro},
  subtopic  = {Behavioral Economics / Altruism},
  type      = {objective},
  statement = {%
    A popular TikTok influencer runs a social experiment called ``Double it
    and give it to the next person.'' The rules are simple: a participant is
    offered \$1 and can either keep it or double it and pass it to the next
    person, who then gets offered \$2. That next person can either keep the
    \$2 or also double it for the following round. The game continues as long
    as participants keep passing.

    Everyone shares the same utility function for this game,
    $u(x, y) = x + \sqrt{200y}$, where $x$ is the participant's own payoff
    and $y$ is the next participant's payoff.

    How many rounds will the game last before a participant decides to keep
    the money instead of passing it?
  },
  choices   = {%
    \item 1 round.
    \item 3 rounds.
    \item 5 rounds.
    \item 7 rounds.
    \item 9 rounds.
  },
  answer    = {E},
  solution  = {%
    At round $n$, the person holds $A_n = 2^{n-1}$ dollars.
    \begin{itemize}
      \item \textbf{Keep:} $x = 2^{n-1}$, $y = 0$~$\Rightarrow$
            $u = 2^{n-1} + 0 = 2^{n-1}$.
      \item \textbf{Pass:} $x = 0$, $y = 2^n$~$\Rightarrow$
            $u = 0 + \sqrt{200 \cdot 2^n}$.
    \end{itemize}

    The person passes if $\sqrt{200 \cdot 2^n} > 2^{n-1}$. Squaring both
    sides:
    \[
      200 \cdot 2^n > 2^{2n-2} \implies 200 > 2^{n-2}
    \]

    Check by round:
    \begin{center}
    \begin{tabular}{ccc}
      \toprule
      Round $n$ & $2^{n-2}$ & Pass? \\
      \midrule
      1  & $0.5$  & Yes \\
      2  & $1$    & Yes \\
      $\vdots$ & $\vdots$ & Yes \\
      9  & $128$  & Yes ($200 > 128$) \\
      10 & $256$  & No  ($200 < 256$) \\
      \bottomrule
    \end{tabular}
    \end{center}

    Participants pass in rounds 1--9 and the 10th person keeps. The game
    lasts \textbf{9 rounds} before someone keeps.
    Answer: \textbf{(E)}.
  },
}
