% q0053.tex — Is this (Deadweight) Loss?
\question{
  id        = {0053},
  title     = {Is this (Deadweight) Loss?},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase C},
  language  = {English},
  topic     = {Micro},
  subtopic  = {Monopoly / Deadweight Loss},
  type      = {objective},
  statement = {%
    A monopolist faces a linear, downward-sloping demand curve and operates
    with constant marginal cost. A technology shock lowers the monopolist's
    marginal cost. What happens to deadweight loss?
  },
  choices   = {%
    \item Deadweight loss increases because the socially optimal quantity
          increases more than the monopoly quantity.
    \item Deadweight loss decreases because firm production approaches the
          social optimum.
    \item Deadweight loss becomes zero because the firm produces the social
          optimum.
    \item Deadweight loss does not change because the fraction of efficient
          quantity produced remains constant.
  },
  answer    = {A},
  solution  = {%
    With linear demand $P = a - bQ$, marginal revenue has twice the slope:
    $MR = a - 2bQ$. When MC falls, the social optimum shifts from
    $P = MC_{\text{old}}$ to $P = MC_{\text{new}}$ (a larger quantity
    increase), while the monopoly output shifts from $MR = MC_{\text{old}}$
    to $MR = MC_{\text{new}}$ (a smaller quantity increase, due to the steeper
    MR slope). The gap between socially optimal and monopoly quantity widens
    absolutely, so deadweight loss increases. (A) is correct.
  },
}
