% q0006.tex — The Good Liar
\question{
  id        = {0006},
  title     = {The Good Liar},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Game Theory},
  subtopic  = {Strategic Behavior / Altruism},
  type      = {objective},
  statement = {%
    Consider a two-player game with the following payoffs. Each player has two
    strategies: Ana can choose Up or Down, while Bob can choose Left or Right.

    \vspace{8pt}
    \begin{center}
    \begin{tabular}{cc|cc|}
      & & \multicolumn{2}{c|}{\textbf{Bob}} \\
      & & \textit{Left} & \textit{Right} \\
      \hline
      \multirow{2}{*}{\textbf{Ana}} & \textit{Up}   & 6,\ 6 & 0,\ 10 \\
                                    & \textit{Down} & 4,\ 0 & 1,\ \ 1 \\
      \hline
    \end{tabular}
    \end{center}
    \vspace{8pt}

    Ana and Bob only care about their own payoffs. However, in a coffee chat
    before the game, Bob was able to fake his intentions and make Ana think he
    is an altruistic person, that is, he equally cares about Ana's payoff as
    his own. What is the outcome of the game if Ana believes Bob's fake
    intentions?
  },
  choices   = {%
    \item (Up, Left).
    \item (Up, Right).
    \item (Down, Left).
    \item (Down, Right).
    \item There is no Nash Equilibrium in this game.
  },
  answer    = {B},
  solution  = {%
    Ana believes Bob has altruistic preferences, so she thinks Bob's utility
    is $U_\text{Bob} = \pi_\text{Bob} + \pi_\text{Ana}$.
    Ana re-evaluates Bob's payoffs accordingly:

    \begin{center}
    \begin{tabular}{c|cc}
            & Left & Right \\
      \hline
      Up    & $6+6=12$ & $10+0=10$ \\
      Down  & $0+4=4$  & $1+1=2$   \\
    \end{tabular}
    \end{center}

    From Ana's perspective (believing Bob is altruistic):
    \begin{itemize}
      \item If Ana plays \textit{Up}: Bob prefers Left ($12 > 10$).
      \item If Ana plays \textit{Down}: Bob prefers Left ($4 > 2$).
    \end{itemize}
    $\Rightarrow$ Bob's dominant strategy (as Ana believes) is \textbf{Left}.

    Knowing Bob will play Left, Ana compares:
    \begin{itemize}
      \item Up + Left: Ana gets $\mathbf{6}$
      \item Down + Left: Ana gets $4$
    \end{itemize}
    $\Rightarrow$ Ana plays \textbf{Up}.

    Outcome: \textbf{(Up, Left)}. Note that in reality Bob would play Right
    (dominant strategy for the real payoffs), but Ana was deceived.
    Answer: \textbf{(A)}.
  },
}
