% q0015.tex — Comparing betas (CAPM)
\question{
  id        = {0015},
  title     = {Comparing Betas},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Finance},
  subtopic  = {CAPM / Asset Pricing},
  type      = {objective},
  statement = {%
    The founders of ObeCorp and ObeconCo started with one idea in the same
    garage. But after disagreeing over who should be CEO, they split up and
    started being competitors in the same industry. You, as a prospective
    investor, are analyzing both companies.

    You already know that the risk-free rate is 3\%, and the expected return
    on the market is 9\%. Additionally, your intern gives you the information
    that ObeCorp has a $\beta$ (systematic risk of assets) of 1.5, while
    ObeconCo has a $\beta$ of 0.8.

    Assuming the Capital Asset Pricing Model (CAPM) holds, what can be said
    about the expected return for each stock?
  },
  choices   = {%
    \item Both companies have the same expected return of 13.5\% since they
          operate in the same industry.
    \item ObeCorp has an expected return of 12\% while ObeconCo has an
          expected return of 8.8\%.
    \item ObeCorp has an expected return of 12\%, because it carries greater
          systematic risk, while ObeconCo has an expected return of 7.8\%.
    \item ObeCorp has an expected return of 13.5\% while ObeconCo has an
          expected return of 7.8\%.
    \item Since the risk-free rate is more relevant than the beta, ObeCorp
          has an expected return of 12\% and ObeconCo of 7\%.
  },
  answer    = {C},
  solution  = {%
    CAPM formula: $E(R_i) = R_f + \beta_i \cdot (E(R_m) - R_f)$

    Market risk premium: $E(R_m) - R_f = 9\% - 3\% = 6\%$

    \begin{itemize}
      \item \textbf{ObeCorp} ($\beta = 1.5$):
            $E(R) = 3\% + 1.5 \times 6\% = 3\% + 9\% = \mathbf{12\%}$
      \item \textbf{ObeconCo} ($\beta = 0.8$):
            $E(R) = 3\% + 0.8 \times 6\% = 3\% + 4.8\% = \mathbf{7.8\%}$
    \end{itemize}

    ObeCorp's higher $\beta$ means it carries greater systematic risk, which
    demands a higher expected return. They operate in the same industry but
    have different risk profiles. Same industry does \emph{not} imply same
    expected return.
    Answer: \textbf{(C)}.
  },
}
