% q0025.tex — Maximizing shareholder value (M&A / IRR) [Numeric Answer]
\question{
  id        = {0025},
  title     = {Maximizing Shareholder Value},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Finance},
  subtopic  = {Corporate Finance / IRR},
  type      = {numeric},
  statement = {%
    An analyst at a manufacturing company is evaluating a vertical M\&A
    opportunity to acquire a key supplier. The company's shareholders are
    risk-neutral. The acquisition would require an upfront payment of
    \$100 million and is expected to generate cash flows in two periods:
    period~1 (short-term) and period~2 (long-term).

    The expected cash flows depend heavily on macroeconomic conditions. The
    analyst has identified three scenarios for the short term:
    \begin{itemize}
      \item \textbf{Base Case} (45\% probability): \$80 million
      \item \textbf{Bear Case} (40\% probability): half of the base case
      \item \textbf{Bull Case} (15\% probability): four times the base case
    \end{itemize}
    The long-term cash flow is expected to be \$200 million across all
    scenarios.

    Given that shareholders are risk-neutral, calculate the Internal Rate of
    Return (IRR) per period for this M\&A project.

    Give the answer as a percentage rounded to the nearest integer. If the
    fractional part is 0.5, round to the nearest even integer.
  },
  choices   = {},
  answer    = {},
  solution  = {%
    % TODO: add solution
    % E[CF1] = 0.45*80 + 0.40*40 + 0.15*320 = 36 + 16 + 48 = 100
    % NPV = -100 + 100/(1+r) + 200/(1+r)^2 = 0
    % Let x = 1/(1+r): 200x^2 + 100x - 100 = 0
    % 2x^2 + x - 1 = 0 => (2x - 1)(x + 1) = 0 => x = 0.5 => r = 100%
  },
}
