% q0059.tex — Vacation, Financed
\question{
  id        = {0059},
  title     = {Vacation, Financed},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase C},
  language  = {English},
  topic     = {Finance},
  subtopic  = {Loan Amortization / Annuity},
  type      = {objective},
  statement = {%
    Raphael borrows \$25{,}000 at a 6\% annual interest rate (compounded
    monthly) to be repaid over 24 monthly instalments. Which of the following
    statements is correct?
  },
  choices   = {%
    \item Doubling the repayment period reduces the monthly payment by half.
    \item If the interest rate doubles, the monthly instalment would also
          double.
    \item If Raphael had brought Leonardo along (total loan \$37{,}500), the
          monthly instalment would rise by 50\%.
    \item If Raphael made more frequent payments, total interest paid would
          stay the same.
  },
  answer    = {C},
  solution  = {%
    The annuity payment formula is:
    \[
      \text{PMT} = PV \cdot \frac{r(1+r)^n}{(1+r)^n - 1},
    \]
    which is \emph{linear} in the principal $PV$. Scaling $PV$ from
    \$25{,}000 to \$37{,}500 is a 50\% increase, so the monthly payment
    rises by exactly 50\%, regardless of the interest rate or term. Options~A,
    B, and~D are incorrect because the non-linear relationship between the
    interest rate/term and PMT means proportional changes in one do not
    produce proportional changes in the other, and more frequent payments
    reduce total interest. (C) is correct.
  },
}
