% q0005.tex — Classicus Met Keynesson (History of Economic Thought)
\question{
  id        = {0005},
  title     = {Classicus Met Keynesson},
  source    = {OBECON},
  year      = {2026},
  round     = {Seletiva IEO -- Phase A},
  language  = {English},
  topic     = {Macro},
  subtopic  = {History of Economic Thought},
  type      = {objective},
  statement = {%
    Consider the following 2020 news report:

    \begin{tcolorbox}[
      colback=white, colframe=black!40, arc=2pt, boxrule=0.5pt,
      left=6pt, right=6pt, top=4pt, bottom=4pt,
    ]
      {\large\bfseries ``Guedes is a liberal, but in a crisis we are all
      Keynesians,'' says Abilio Diniz}\\[4pt]
      {\small\textit{Brazilian businessman explained at conference that Economy
      Minister should increase stimulus packages against coronavirus}}
    \end{tcolorbox}

    \vspace{6pt}
    The irony presented by the news is the opposition between the liberal model
    and the Keynesian model for solving economic crises.

    During the COVID-19 pandemic in 2020, the economies of several countries
    slowed down sharply. Suppose that two economists had been hired by the
    Brazilian government at that time to design mechanisms to stimulate the
    country's economy: Classicus, who follows the classical economic tradition,
    and Keynesson, who follows the Keynesian tradition. Knowing that both aim
    to rebalance the level of goods and services produced in the country (GDP),
    that is, to achieve the level of full employment, mark the correct
    alternative.
  },
  choices   = {%
    \item Keynesson proposes an increase in government spending to stimulate
          the economy through aggregate demand and, in this way, raise the
          level of output.
    \item For Classicus, the solution to the crisis is an increase in the
          money supply, which will lead to a fall in the interest rate and a
          consequent rise in aggregate demand, stimulating the economy and
          increasing output.
    \item For Keynesson, the only possible solution to the crisis would be a
          reduction in the country's tax burden, which would increase
          households' disposable income and stimulate consumption and output;
          other mechanisms, such as higher government spending, would have no
          effect.
    \item Both Classicus and Keynesson agree that an increase in government
          spending raises aggregate demand and, consequently, output, although
          they disagree about the magnitude of the resulting inflationary
          effect.
    \item For Keynesson, an increase in coronavirus stimulus packages raises
          government spending, which consequently increases aggregate supply
          and rekindles the economy, increasing output.
  },
  answer    = {},
  solution  = {%
    % TODO: add solution
  },
}
